---
title: pvaluefunctions
output:
  github_document:
    toc: true
    toc_depth: 3
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, 
  comment = "#>",
  fig.path = "man/figures/README-"
)
```

# *P*-value functions <img src="man/figures/logo2.svg" align="right" width="120" />

[![CRAN\_Status\_Badge](http://www.r-pkg.org/badges/version/pvaluefunctions)](https://cran.r-project.org/package=pvaluefunctions)
[![downloads](https://cranlogs.r-pkg.org/badges/pvaluefunctions)](https://cran.r-project.org/package=pvaluefunctions)
[![total downloads](https://cranlogs.r-pkg.org/badges/grand-total/pvaluefunctions)](http://cranlogs.r-pkg.org/badges/grand-total/pvaluefunctions)
[![Rdoc](http://www.rdocumentation.org/badges/version/pvaluefunctions)](http://www.rdocumentation.org/packages/pvaluefunctions)

## Accompanying paper

We published an [accompanying paper](https://doi.org/10.1002/sim.8293) to illustrate the use of *p*-value functions:

Infanger D, Schmidt-Trucksäss A. (2019): *P* value functions: An underused method to present research results and to promote quantitative reasoning. *Statistics in Medicine,* 1-9. doi: 10.1002/sim.8293.

### Recreation of the figures in the paper

The code and instructions to reproduce all graphics in our paper can be found in the following GitHub repository: https://github.com/DInfanger/pvalue_functions

## Overview

This is the repository for the R-package [`pvaluefunctions`](https://cran.r-project.org/package=pvaluefunctions). The package contains R functions to create graphics of *p*-value functions, confidence distributions, confidence densities, or the [Surprisal value (S-value)](http://www.umsl.edu/~fraundorfp/egsurpri.html) (Greenland 2019).

## Installation

You can install the package directly from CRAN by typing `install.packages("pvaluefunctions")`. After installation, load it in R using `library(pvaluefunctions)`.

```{r source_github, message = FALSE, warning = FALSE, echo = FALSE, eval = FALSE}
library(devtools)

# Load main function
source_url("https://raw.githubusercontent.com/DInfanger/pvaluefunctions/master/R/confidence_distributions.R")

```

## Dependencies

The function depends on the following R packages, which need to be installed beforehand:

* ggplot2
* scales
* zipfR

Use the command `install.packages(c("ggplot2", "scales", "zipfR"))` in R to install those packages.

### Important information!

The newest version of [ggplot2 (3.1.1)](https://CRAN.R-project.org/package=ggplot2) has a [bug](https://github.com/tidyverse/ggplot2/issues/2978) in `sec_axis` that will lead to the secondary y-axis being labelled wrongly.

It is therefore recommended that you install the developmental version of ggplot2 until the bug has been fixed. You can install the developmental version using the following command (after installing the [`devtools`](https://CRAN.R-project.org/package=devtools) package): `devtools::install_github("tidyverse/ggplot2")`

To see what version of ggplot2 has been used to create the plots on this page, see the [Session info].

This warning will be deleted upon the release of a new version of ggplot2 that fixes the bug.

## Examples

For more examples and code, see the [vignette](https://CRAN.R-project.org/package=pvaluefunctions/vignettes/pvaluefun.html).

```{r ttest_pval, echo = FALSE, message = FALSE, warning = FALSE, fig.width = 9, fig.height = 7, out.width = "70%", fig.align='left', dev = "png", dev.args = list(type = "cairo-png"), dpi = 600}

devtools::install_github("DInfanger/pvaluefunctions")

library(pvaluefunctions)

#-----------------------------------------------------------------------------
# Create p-value function
#-----------------------------------------------------------------------------

res <- conf_dist(
  estimate = c(-1.58)
  , df = c(17.77647)
  , tstat = c(-1.860813)
  , type = "ttest"
  , plot_type = "p_val"
  , n_values = 1e4L
  # , est_names = c("")
  , conf_level = c(0.95, 0.90, 0.80)
  , null_values = c(0)
  , trans = "identity"
  , alternative = "two_sided"
  , log_yaxis = FALSE
  , cut_logyaxis = 0.05
  , xlab = "Mean difference (group 1 - group 2)"
  , together = FALSE
  , plot_p_limit = 1 - 0.999
  , plot_counternull = TRUE
  , inverted = FALSE
  , title = expression(italic(P)~"-value function for a"~italic(t)~"-test")
)
```

```{r ttest_sval, echo = FALSE, message = FALSE, warning = FALSE, fig.width = 9, fig.height = 7, out.width = "70%", fig.align='left', dev = "png", dev.args = list(type = "cairo-png"), dpi = 600}

#-----------------------------------------------------------------------------
# Create s-value function
#-----------------------------------------------------------------------------

res <- conf_dist(
  estimate = c(-1.58)
  , df = c(17.77647)
  , tstat = c(-1.860813)
  , type = "ttest"
  , plot_type = "s_val"
  , n_values = 1e4L
  # , est_names = c("")
  , conf_level = c(0.95, 0.90, 0.80)
  , null_values = c(0)
  , trans = "identity"
  , alternative = "two_sided"
  , log_yaxis = FALSE
  , cut_logyaxis = 0.05
  , xlab = "Mean difference (group 1 - group 2)"
  , together = FALSE
  , plot_p_limit = 1 - 0.999
  , plot_counternull = TRUE
  , inverted = FALSE
  , title = expression(italic(S)~"-value function for a"~italic(t)~"-test")
)
```


```{r benderfig1, echo = FALSE, message = FALSE, warning = FALSE, fig.width = 9, fig.height = 7, out.width = "70%", fig.align='left', dev = "png", dev.args = list(type = "cairo-png"), dpi = 600}
rse_fun <- function(x){
  100*(1 - exp(x))
}

rse_fun_inv <- function(x){
  log(1 - (x/100))
}

res <- conf_dist(
  estimate = c(log(0.72))
  , stderr = (0.187618)
  , type = "coxreg"
  , plot_type = "p_val"
  , n_values = 1e4L
  , est_names = c("RSE")
  , conf_level = c(0.95, 0.8, 0.5)
  , null_values = rse_fun_inv(c(0))
  , trans = "rse_fun"
  , alternative = "two_sided"
  , log_yaxis = FALSE
  , cut_logyaxis = 0.05
  , xlab = "Relative survival effect (1 - HR%)"
  , xlim = rse_fun_inv(c(-30, 60))
  , together = FALSE
  , plot_p_limit = 1 - 0.999
  , plot_counternull = TRUE
  , inverted = TRUE
  , title = "Figure 1 in Bender et al. (2005)"
)
```

## References

Bender R, Berg G, Zeeb H. (2005): Tutorial: using confidence curves in medical research. *Biom J.* 47(2): 237-47.

Fraser  D. A. S. (2019): The *p*-value function and statistical inference. *Am Stat,* 73:sup1, 135-147.

Greenland S (2019): Valid *P*-Values Behave Exactly as They Should: Some Misleading Criticisms of *P*-Values and Their Resolution with *S*-Values. *Am Stat,* 73sup1, 106-114.

Infanger D, Schmidt-Trucksäss A. (2019): *P* value functions: An underused method to present research results and to promote quantitative reasoning. *Stat Med,* 1-9. doi: 10.1002/sim.8293.

Poole C. (1987a): Beyond the confidence interval. *Am J Public Health.* 77(2): 195-9.

Poole C. (1987b) Confidence intervals exclude nothing. *Am J Public Health.* 77(4): 492-3.

Rosenthal R, Rubin DB. (1994): The counternull value of an effect size: A new statistic. Psychol Sci. 5(6): 329-34.

Schweder T, Hjort NL. (2016): Confidence, likelihood, probability: statistical inference with confidence distributions. New York, NY: Cambridge University Press.

Xie M, Singh K, Strawderman WE. (2011): Confidence Distributions and a Unifying Framework for Meta-Analysis. *J Am Stat Assoc* 106(493): 320-33. doi: 10.1198/jasa.2011.tm09803.

Xie Mg, Singh K. (2013): Confidence distribution, the frequentist distribution estimator of a parameter: A review. *Internat Statist Rev.* 81(1): 3-39.

## Contact

[Denis Infanger](https://dsbg.unibas.ch/de/personen/denis-infanger/)

## Session info

```{r session_info, include=TRUE, echo=FALSE}
sessionInfo()
```

## License

[![License: GPL v3](https://img.shields.io/badge/License-GPL%20v3-blue.svg)](https://www.gnu.org/licenses/gpl-3.0)
