---
title: "*P*-value functions"
author: "Denis Infanger"
date: "`r Sys.Date()`"
output: html_vignette
vignette: >
  %\VignetteIndexEntry{*P*-value functions}
  %\VignetteEngine{knitr::knitr}
  \usepackage[utf8]{inputenc}
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE
  , comment = "#>"
  # , tidy.opts = list(width.cutoff = 160, tidy = FALSE)
)
```

## Overview

This package contains R functions to create graphics of *p*-value functions, confidence distributions, confidence densities, or the [Surprisal value (S-value)](http://www.umsl.edu/~fraundorfp/egsurpri.html) (Greenland 2019). An R-script to reproduce the plots in the publication is also available.

## Installation

Download the file `confidence_distributions.R` to your computer. You can either `source()` the function in R or open it, select and run everything. After loading the function, it's ready for use.

To reproduce the plots from the publication, download the file `paper_plots.R` and run it *after* loading the main function contained in the file `confidence_distributions.R` (see above).

Alternatively, you can source the files directly from the GitHub repository using the [`devtools`](https://CRAN.R-project.org/package=devtools) package:

```{r source_github, message = FALSE, warning = FALSE, echo = TRUE, eval = TRUE}
library(devtools)

# Load function
source_url("https://raw.githubusercontent.com/DInfanger/pvalue_functions/master/confidence_distributions.R")

```


## Dependencies

The function depends on the following R packages, which need to be installed beforehand:

* ggplot2
* scales
* zipfR

Use the command `install.packages(c("ggplot2", "scales", "zipfR"))` in R to install those packages.

### Important information!

The newest version of [ggplot2 (3.1.1)](https://CRAN.R-project.org/package=ggplot2) has a [bug](https://github.com/tidyverse/ggplot2/issues/2978) in `sec_axis` that will lead to the secondary y-axis being labelled wrongly.

It is therefore recommended that you install the developmental version of ggplot2 until the bug has been fixed. You can install the developmental version using the following command (after installing the [`devtools`](https://CRAN.R-project.org/package=devtools) package): `devtools::install_github("tidyverse/ggplot2")`

## Usage

There is only one function needed to create the plots: `conf_dist()`. The function has the following arguments:

* **`estimate`**: Numerical vector containing the estimate(s).
* **`n`**: Numerical vector containing the sample size(s). Required for correlations, variances and proportions. Must be equal the number of estimates.
* **`df`**: Numerical vector containing the degrees of freedom. Required for statistics based on the *t*-distribution (e.g. linear regression) and *t*-tests. Must be equal the number of estimates.
* **`stderr`**: Numerical vector containing the standard error(s) of the estimate(s). Required for statistics based on the *t*-distribution (e.g. linear regression) and the normal distribution (e.g. logistic regression). Must be equal the number of estimate(s).
* **`tstat`**: Numerical vector containing the *t*-statistic(s). Required for *t*-tests (means and mean differences). Must be equal the number of estimates. 
* **`type`**: String indicating the type estimate. Must be one of the following: `ttest`, `linreg`, `gammareg`, `general_t`, `logreg`, `poisreg`, `coxreg`, `general_z`, `pearson`, `spearman`, `kendall`, `var`, `prop`.
* **`plot_type`**: String indicating the type of plot. Must be one of the following: `cdf` (confidence distribution), `pdf` (confidence density), `p_val` (*p*-value function), `s_val` (Surprisal).
* **`n_values`** (optional): Integer indicating the number of points that are used to generate the graphics. The higher this number, the higher the computation time and resolution.
* **`est_names`** (optional): String vector indicating the names of the estimate(s). Must be equal the number of estimates.
* **`conf_level`** (optional): Numerical vector indicating the confidence level(s). Bust be between 0 and 1.
* **`null_values`** (optional): Numerical vector indicating the null value(s) in the plot
* **`trans`** (optional): String indicating the transformation function that will be applied to the estimates and confidence curves. For example: "exp" for an exponential transformation of the log-odds in logistic regression. 
* **`alternative`**: String indicating if the confidence level(s) are two-sided or one-sided. Must be one of the following: `two_sided`, `one_sided`.
* **`log_yaxis`**: Logical. Indicating if a portion of the y-axis should be displayed on the logarithmic scale.
* **`cut_logyaxis`**: Numerical value indicating the threshold below which the y-axis will be displayed logarithmically. Must lie between 0 and 1.
* **`xlab`** (optional): String indicating the label of the x-axis.
* **`xlim`** (optional): Optional numerical vector of length 2 indicating the limits of the x-axis on the *untransformed* scale.
* **`together`**: Logical. Indicating if graphics for multiple estimates should be displayed together or on separate plots.
* **`plot_p_limit`**: Numerical value indicating the lower limit of the y-axis. Must be greater than 0 for a logarithmic scale (i.e. `log_yaxis = TRUE`).

### Required arguments for different estimate types

* $t$-tests: `estimate`, `df`, `tstat`.
* Linear regression, Gamma regression, general estimates based on the *t*-distribution: `estimate`, `df`, `stderr`.
* Logistic regression, Poisson regression, Cox regression, general estimates based on the normal distribution: `estimate`, `stderr`.
* Correlation coefficients (Pearson, Spearman, Kendall), proportions, variances: `estimate`, `n`.

### Returned values

The main function `conf_dist()` returns five objects in a list:

* **`res_frame`**: A data frame containing the values used to construct the plot.
* **`conf_frame`**: A data frame containing the confidence intervals for the specified confidence levels for all estimates.
* **`counternull_frame`**: A data frame containing the counternull values for the specified null values (see Rosenthal & Rubin (1994) for more information about the counternull).
* **`point_est`**: A data frame containing the point estimates for all estimates. The point estimates correspond to the mean, median or mode of the confidence density (see Xie & Singh (2013) for more information). Estimates are produced using numerical procedures: Increase the number of points `n_values` for higher numerical precision.
* **`plot`**: A [ggplot2](https://ggplot2.tidyverse.org/) plot object.

## Examples

### Two-sample *t*-test with unequal variances (Welch-Test)

```{r ttest, message = FALSE, warning = FALSE, fig.width = 9, fig.height = 7, out.width = "80%", fig.align='center', dev = "png", dev.args = list(type = "cairo-png"), dpi = 600}

#-----------------------------------------------------------------------------
# Sourcing function
#-----------------------------------------------------------------------------

#source("confidence_distributions.R")

#-----------------------------------------------------------------------------
# T-Test
#-----------------------------------------------------------------------------

with(sleep, mean(extra[group == 1])) - with(sleep, mean(extra[group == 2]))
t.test(extra ~ group, data = sleep, var.equal = FALSE)

#-----------------------------------------------------------------------------
# Create p-value function
#-----------------------------------------------------------------------------

res <- conf_dist(
  estimate = c(-1.58)
  , df = c(17.77647)
  , tstat = c(-1.860813)
  , type = "ttest"
  , plot_type = "p_val"
  , n_values = 1e4L
  # , est_names = c("")
  , conf_level = c(0.95, 0.90, 0.80)
  , null_values = c(0)
  , trans = "identity"
  , alternative = "two_sided"
  , log_yaxis = TRUE
  , cut_logyaxis = 0.05
  , xlab = "Mean difference (group 1 - group 2)"
  , together = FALSE
  , plot_p_limit = 1 - 0.999
)
```

### Single coefficient from a linear regression model
#### *P*-value function

```{r linreg_single_pval, message = FALSE, warning = FALSE, fig.width = 9, fig.height = 7, out.width = "80%", fig.align='center', dev = "png", dev.args = list(type = "cairo-png"), dpi = 600}
#-----------------------------------------------------------------------------
# Model
#-----------------------------------------------------------------------------

mod <- lm(Infant.Mortality~Agriculture + Fertility + Examination, data = swiss)

summary(mod)

#-----------------------------------------------------------------------------
# Create p-value function
#-----------------------------------------------------------------------------

res <- conf_dist(
  estimate = c(-0.02143)
  , df = c(43)
  , stderr = (0.02394)
  , type = "linreg"
  , plot_type = "p_val"
  , n_values = 1e4L
  # , est_names = c("")
  , conf_level = c(0.95, 0.90, 0.80)
  , null_values = c(0)
  , trans = "identity"
  , alternative = "two_sided"
  , log_yaxis = TRUE
  , cut_logyaxis = 0.05
  , xlab = "Coefficient Agriculture"
  , together = FALSE
  , plot_p_limit = 1 - 0.999
)
```

#### Confidence distribution
```{r linreg_single_cdf, message = FALSE, warning = FALSE, fig.width = 9, fig.height = 7, out.width = "80%", fig.align='center', dev = "png", dev.args = list(type = "cairo-png"), dpi = 600}
res <- conf_dist(
  estimate = c(-0.02143)
  , df = c(43)
  , stderr = (0.02394)
  , type = "linreg"
  , plot_type = "cdf"
  , n_values = 1e4L
  # , est_names = c("")
  , conf_level = c(0.95, 0.90, 0.80)
  , null_values = c(0)
  , trans = "identity"
  , alternative = "two_sided"
  # , log_yaxis = TRUE
  # , cut_logyaxis = 0.05
  , xlab = "Coefficient Agriculture"
  , xlim = c(-0.12, 0.065)
  , together = FALSE
  # , plot_p_limit = 1 - 0.999
)

```


### Multiple coefficients from a linear regression model
#### *P*-value functions

```{r linreg_multiple_pval, message = FALSE, warning = FALSE, fig.width = 9, fig.height = 7, out.width = "80%", fig.align='center', dev = "png", dev.args = list(type = "cairo-png"), dpi = 600}
res <- conf_dist(
  estimate = c(0.13115, 0.04913)
  , df = c(43, 43)
  , stderr = c(0.04145, 0.08351)
  , type = "linreg"
  , plot_type = "p_val"
  , n_values = 1e4L
  , est_names = c("Fertility", "Examination")
  , conf_level = c(0.95, 0.90, 0.80)
  , null_values = c(0)
  , trans = "identity"
  , alternative = "two_sided"
  , log_yaxis = TRUE
  , cut_logyaxis = 0.05
  , xlab = "Coefficients"
  , together = TRUE
  , plot_p_limit = 1 - 0.999
)
```

#### Surprisal values
```{r linreg_multiple_sval, message = FALSE, warning = FALSE, fig.width = 9, fig.height = 7, out.width = "80%", fig.align='center', dev = "png", dev.args = list(type = "cairo-png"), dpi = 600}
res <- conf_dist(
  estimate = c(0.13115, 0.04913)
  , df = c(43, 43)
  , stderr = c(0.04145, 0.08351)
  , type = "linreg"
  , plot_type = "s_val"
  , n_values = 1e4L
  , est_names = c("Fertility", "Examination")
  , conf_level = c(0.95, 0.90, 0.80)
  , null_values = c(0)
  , trans = "identity"
  , alternative = "two_sided"
  # , log_yaxis = TRUE
  # , cut_logyaxis = 0.05
  , xlab = "Coefficients"
  , together = TRUE
  , plot_p_limit = 1 - 0.999
)
```

### Pearson correlation coefficient (one-sided)

```{r corr_pearson, message = FALSE, warning = FALSE, fig.width = 9, fig.height = 7, out.width = "80%", fig.align='center', dev = "png", dev.args = list(type = "cairo-png"), dpi = 600}
#-----------------------------------------------------------------------------
# Calculate Pearson's correlation coefficient
#-----------------------------------------------------------------------------

cor.test(swiss$Fertility, swiss$Agriculture, alternative = "two.sided", method = "pearson")

#-----------------------------------------------------------------------------
# Create p-value function
#-----------------------------------------------------------------------------

res <- conf_dist(
  estimate = c(0.3530792)
  , n = 47
  , type = "pearson"
  , plot_type = "p_val"
  , n_values = 1e4L
  # , est_names = c("")
  , conf_level = c(0.95, 0.90, 0.80)
  , null_values = c(0)
  , trans = "identity"
  , alternative = "one_sided"
  , log_yaxis = TRUE
  , cut_logyaxis = 0.05
  , xlab = "Pearson correlation"
  , together = TRUE
  , plot_p_limit = 1 - 0.999
)
```

### Odds ratio from logistic regression

```{r logreg, message = FALSE, warning = FALSE, fig.width = 9.2, fig.height = 7.2, out.width = "80%", fig.align='center', dev = "png", dev.args = list(type = "cairo-png"), dpi = 600}
#-----------------------------------------------------------------------------
# Calculate logistic regression model using a dataset from UCLA
#-----------------------------------------------------------------------------

dat_tmp <- read.csv("https://stats.idre.ucla.edu/stat/data/binary.csv")

dat_tmp$rank <- factor(dat_tmp$rank)
logistic_mod <- glm(admit ~ gre + gpa + rank, data = dat_tmp, family = "binomial")

summary(logistic_mod)

rm(dat_tmp)

#-----------------------------------------------------------------------------
# Create p-value function
#-----------------------------------------------------------------------------

res <- conf_dist(
  estimate = c(0.804037549)
  , stderr = c(0.331819298)
  , type = "logreg"
  , plot_type = "p_val"
  , n_values = 1e4L
  , est_names = c("GPA")
  , conf_level = c(0.95, 0.90, 0.80)
  , null_values = c(log(1))
  , trans = "exp"
  , alternative = "two_sided"
  , log_yaxis = TRUE
  , cut_logyaxis = 0.05
  , xlab = "Odds Ratio (GPA)"
  , xlim = log(c(0.4, 5))
  , together = FALSE
  , plot_p_limit = 1 - 0.999
)
```

### Proportion

```{r prop, message = FALSE, warning = FALSE, fig.width = 9, fig.height = 7, out.width = "80%", fig.align='center', dev = "png", dev.args = list(type = "cairo-png"), dpi = 600}
res <- conf_dist(
  estimate = c(0.44)
  , n = c(50)
  , type = "prop"
  , plot_type = "p_val"
  , n_values = 1e4L
  # , est_names = c("")
  , conf_level = c(0.95, 0.90, 0.80)
  , null_values = c(0.5)
  , trans = "identity"
  , alternative = "two_sided"
  , log_yaxis = TRUE
  , cut_logyaxis = 0.05
  , xlab = "Proportion"
  # , xlim = log(c(0.95, 1.2))
  , together = FALSE
  , plot_p_limit = 1 - 0.999
)
```

## References

Bender R, Berg G, Zeeb H. (2005): Tutorial: using confidence curves in medical research. *Biom J.* 47(2): 237-47.

Fraser  D. A. S. (2019): The *p*-value function and statistical inference. *The American Statistician,* 73:sup1, 135-147.

Greenland S (2019): Valid *P*-Values Behave Exactly as They Should: Some Misleading Criticisms of *P*-Values and Their Resolution with *S*-Values. *The American Statistician,* 73sup1, 106-114.

Poole C. (1987a): Beyond the confidence interval. *Am J Public Health.* 77(2): 195-9.

Poole C. (1987b) Confidence intervals exclude nothing. *Am J Public Health.* 77(4): 492-3.

Rosenthal R, Rubin DB. (1994): The counternull value of an effect size: A new statistic. Psychol Sci. 5(6): 329-34.

Schweder T, Hjort NL. (2016): Confidence, likelihood, probability: statistical inference with confidence distributions. New York, NY: Cambridge University Press.

Xie M, Singh K, Strawderman WE. (2011): Confidence Distributions and a Unifying Framework for Meta-Analysis. *J Am Stat Assoc* 106(493): 320-33. doi: 10.1198/jasa.2011.tm09803.

Xie Mg, Singh K. (2013): Confidence distribution, the frequentist distribution estimator of a parameter: A review. *Internat Statist Rev.* 81(1): 3-39.

## Contact

[Denis Infanger](https://dsbg.unibas.ch/de/personen/denis-infanger/)

## Session info

```{r session_info, include=TRUE, echo=FALSE}
sessionInfo()
```

## License

[![License: GPL v3](https://img.shields.io/badge/License-GPL%20v3-blue.svg)](https://www.gnu.org/licenses/gpl-3.0)

[GNU General Public License v3.0](https://www.gnu.org/licenses/gpl-3.0).













